Hydra Battles and Cie

### 2021-01-18 doc/chapter-primrec.tex
redaction of Ackermann chapter (continued)

\begin{remark}
Beware of the conventions used in the \texttt{primRec} library!
The constructor (\texttt{projFunc $n$ $m$})  is associated with the projection $\pi_{n-m,n}$ and \emph{not}
$\pi_{n, m}$.
For instance, the projection $\pi_{2,5}$ defined by $\pi_{2,5}(a,b,c,d,e)=b$ corresponds to the term
(\texttt{projFunc 5 3 H}), where \texttt{H} is a proof of $3<5$.
 This fact is reported in the comments of \texttt{primRec.v}. We presume that this convention makes it easier to define the evaluation function \texttt{evalProjFunc $n$} (see next sub-section) . Trying the other conention is left as an exercise.
\end{remark}

@@

>>>>

\begin{remark}
\label{projFunc-order-of-args}
Beware of the conventions used in the \texttt{primRec} library!
The constructor (\texttt{projFunc $n$ $m$})  is associated with the projection $\pi_{n-m,n}$ and \emph{not}
$\pi_{n, m}$.
For instance, the projection $\pi_{2,5}$ defined by $\pi_{2,5}(a,b,c,d,e)=b$ corresponds to the term
(\texttt{projFunc 5 3 H}), where \texttt{H} is a proof of $3<5$.
 This fact is reported in the comments of \texttt{primRec.v}. We presume that this convention makes it easier to define the evaluation function \texttt{evalProjFunc $n$} (see next sub-section) . Trying the other conention is left as an exercise.
\end{remark}

@@

\subsubsection{Elementary Proofs of \texttt{isPR} statements}

The constructors \texttt{zeroFunc}, \texttt{succFunc},  and \texttt{projFunc} of type
\texttt{PrimRec} allows us to write trivial proofs of primitive recursivity. 
Although  the following lemmas are alreday proven in 
\href{../theories/html/hydras.Ackermann.primRec.html}{Ackermann.primRec.v},
we wrote alternate proofs in 
\href{../theories/html/hydras.Ackermann.PrimRecExamples.html}%
{Ackermann.PrimRecExamples.v}, in order to illustrate the main proof patterns.

\begin{Coqsrc}
Module Alt.

Lemma zeroIsPR : isPR 0 0.
Proof.
  exists zeroFunc.
\end{Coqsrc}

\begin{Coqanswer}
1 subgoal (ID 90)

  ============================
  extEqual 0 (evalPrimRec 0 zeroFunc) 0
\end{Coqanswer}

\begin{Coqanswer}
1 subgoal (ID 91)

  ============================
  0 = 0
\end{Coqanswer}

\begin{Coqsrc}
  reflexivity.
Qed.
\end{Coqsrc}

Likewise, we prove that the successor function on \texttt{nat} is primitive recursive too.

\begin{Coqsrc}
Lemma SuccIsPR : isPR 1 S.
Proof.
  exists succFunc; cbn; reflexivity.
Qed.
\end{Coqsrc}

Projections are proved primitive recursive, case by case (many examples in 
\href{../theories/html/hydras.Ackermann.primRec.html}{Ackermann.primRec.v}).
\emph{Please notice again that the name of the projection follows the mathematical tradition, 
whilst the arguments of  \texttt{projFunc} use another convention (\emph{cf} remark~\vref{projFunc-order-of-args}).}

\begin{Coqsrc}
Lemma pi2_5IsPR : isPR 5 (fun a b c d e => b).
Proof.
 assert (H: 3 < 5) by auto.
 exists (projFunc 5 3 H).
 cbn; reflexivity.
Qed.
\end{Coqsrc}

\subsubsection{Using Function Composition}

Let us look at the proof that any constant $n$ of type \texttt{nat} has type (\texttt{PR 0})
(lemma  \texttt{const1\_NIsPR} of \texttt{primRec}). We carry out a proof by induction on $n$, the base case of which is already proven.
Now, let us assume $n$ is \texttt{PR $n$}, with $x:\texttt{PrimRec}\,0$ as a ``realizer''.
Thus we would like to compose this constant function with the unary successor function.

This is exactly the role of the instance \texttt{composeFunc 0 1} of the dependently typed
function \texttt{composeFunc}, as shown by the following lemma.

\begin{Coqsrc}
Fact compose_01 :
    forall (x:PrimRec 0) (t : PrimRec 1),
    let c := evalPrimRec 0 x in
    let f := evalPrimRec 1 t in
    evalPrimRec 0 (composeFunc 0 1
                               (PRcons 0 0 x (PRnil 0))
                               t)  =
     f c.
Proof. reflexivity. Qed.
\end{Coqsrc}

Thus, we get a quite simple proof of \texttt{const1\_NIsPR}.

\vspace{4pt}
\noindent
\emph{From\href{../theories/html/hydras.Ackermann.PrimRecExamples.html}{Ackermann.PrimRecExamples}}.
\begin{Coqsrc}
Lemma  const1_NIsPR n : isPR 0 n. 
Proof.
  induction n.
 - apply zeroIsPR.
 - destruct IHn as [x Hx].
   exists (composeFunc 0 1 (PRcons 0 0 x (PRnil 0)) succFunc). 
   cbn in *; intros; now rewrite Hx.
Qed.
\end{Coqsrc}
